Titeln på avhandlingens språk

### 2023-04-22 body/part1/chapter2-relatedwork/key_contributions.tex
contrib tree

This thesis presents two key contributions to literature.
First, we introduce and study a methodology for the performance evaluation of latency-sensitive, feedback-loop-based applications enabled by Edge Computing such as \gls{MAR} and \glspl{CPS}.
Our approach is based on the novel idea of emulating \emph{only} the end-user component of such applications while using the \emph{real} network and compute components of the system.
In other words, our methodology involves replacing the client side of the system with commercial off-the-shelf (COTS) general-purpose computing devices running realistic software imitations of the desired behaviors, while retaining the real network and compute hardware at the backend.
This entails modeling and emulating human inputs in \gls{MAR} applications and the physical components of \glspl{CPS}.

>>>>

\begin{figure}
    \centering
    \begin{forest}
        for tree={% style of tree
            align=center,
            grow=0,reversed, % tree direction
            parent anchor=east,
            child anchor=west, % edge anchors
            edge={line cap=round},
            outer sep=+1pt, % edge/node connection
            rounded corners,minimum width=15mm,minimum height=8mm, % node shape
            l sep=7mm % level distance
        }
        [{\scshape Contributions}
            [Methodology
                [{Use-case\\studies} [\acs{WCA}] [\acsp{NCS}]]
                [{Tooling} [{EdgeDroid\\(\acs{WCA})}] [{\acs{CLEAVE}\\(\acsp{NCS})}] [Ainur\\(testbed)]]
            ]
            [{Implications\\of Realism}
                [{Characterization of\\human behavior in \acs{WCA}}]
                [{Realistic human\\model for \acs{WCA}}]
                [Exploration of avenues\\for optimization of \acs{WCA}]
            ]
        ]
        \end{forest}
    \caption{Hierarchical overview of the main and secondary contributions of this dissertation.}\label{fig:contribs}
\end{figure}

This thesis presents two key contributions and a number of sub-contributions to literature, as illustrated in \cref{fig:contribs}.
Firstly, we introduce a methodology for the performance evaluation of latency-sensitive, feedback-loop-based applications enabled by Edge Computing such as \gls{MAR} and \glspl{CPS}.
This methodology is validated by example through two representative use-cases, and we provide tooling for its implementation.
Secondly, we study the implications improved client-side emulation realism has on the results produced with our methodology.
These implications are again explored by example.
We design and implement of a detailed model of human user behavior in \gls{WCA}, and then use it to study the potential for optimization of these systems.

Our first contribution corresponds to a methodological approach to the study and performance evaluation of Edge-enabled latency-sensitive applications. 
This approach is based on the novel idea of emulating \emph{only} the end-user component of such applications while using the \emph{real} network and compute components of the system.
In other words, our methodology involves replacing the client side of the system with commercial off-the-shelf (COTS) general-purpose computing devices running realistic software imitations of the desired behaviors, while retaining the real network and compute hardware at the backend.
This entails modeling and emulating human inputs in \gls{MAR} applications and the physical components of \glspl{CPS}.

### 2023-04-22 body/part1/chapter3-contributions/methodology.tex
contrib tree

As summarized in \cref{chap:introduction}, this thesis presents a comprehensive study on latency-sensitive applications in Edge Computing.
The primary contribution of this research is a methodology for analyzing and evaluating the performance of these applications.
The methodology is validated through the examination of two initial case studies, \gls{WCA} and \glspl{NCS}.
In addition to the methodology, this thesis includes several other key contributions.
A prototype testbed and accompanying software framework have been developed to provide a platform for conducting experiments in Edge Computing.
Furthermore, a further deeper exploration into the \gls{WCA} case studied yielded the first-ever model of human behavior in \gls{WCA}, which is a significant advancement in the field.
The methodology, the case studies, and the prototype testbed provide valuable insights into the performance of latency-sensitive applications in Edge Computing and can be used as a basis for future research in this area.

@@

This thesis makes a contribution to the field of Edge Computing by investigating the applications of a methodology for the study of latency-sensitive applications deployed on Edge infrastructure.
These applications, such as \gls{WCA} and \glspl{NCS}, are challenging to study and scale due to their cyber-physical nature, complexity, and requirements for low latency and high reliability.

The main focus of this work is to enhance the accuracy and realism of results related to Edge infrastructure, particularly in regard to the network.
The proposed methodology aims to provide a more comprehensive and realistic assessment of the performance of these latency-sensitive applications, allowing for a better understanding of the strengths and limitations of Edge Computing in this context.
Our results will contribute to the development of new techniques and approaches for improving the performance and reliability of Edge Computing systems.

@@

% As discussed in \cref{chap:relatedwork}, most research in this field falls into either simulated or fully-practical real-world approaches.
Our emulated methodology presents several advantages over alternatives, simulated or fully-practical approaches.
To begin with, and perhaps most importantly, it makes our approach portable and applicable to real Edge computing deployments.
By limiting the emulation to the workload, our approach can be directly employed on pre-existing Edge computing infrastructure in order to benchmark applications before they are deployed.
This makes our approach valuable to not only to the research community, but also to system designers and application developers.

@@

\medskip
In the following subsections we will discuss two illustrative case studies for our methodology; \acl{WCA} and \acl{NCS} research and benchmarking.
\Cref{summary:methodology:usecase_wca} discusses the application of our methodology to \gls{WCA}.
This work was originally published in \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid}.
Next, in \cref{summary:methodology:usecase_ncs} we expand this discussion to \glspl{NCS}, a discussion originally approached in \cref{paper:olguinmunoz2022cleave}.

\subsection{Case study: \acl{WCA}}\label{summary:methodology:usecase_wca}

@@

\subsection{Case study: \acsp{NCS}}\label{summary:methodology:usecase_ncs}

We discuss an initial case study of our methodology in the context of \glspl{NCS} and similar \glspl{CPS} in \cref{paper:olguinmunoz2022cleave}.
In this work, we present \gls{CLEAVE}, a software-based framework for scalable and repeatable benchmarking of Edge-native \aclp{NCS}.
The growing adoption of Edge computing has made it important to understand the strengths and weaknesses of different Edge concepts for the deployment of such \glspl{CPS}.

@@

After our initial implementation and experimentation with our methodology for \gls{WCA} applications, we realized that creating a testbed for experimental Edge Computing research for would significantly enhance our work.
Establishing such testbed would provide a dedicated and controlled environment that can accelerate future research by eliminating the need for ad-hoc implementations for each new study.
Moreover, having a standardized baseline would make it easier to compare results across different studies.
The testbed could serve as a platform for researchers to test new ideas and compare them to the results of previous experiments.
This would enable more efficient and accurate assessments of the methodology's performance and facilitate the development of new methods and techniques.
Additionally, by establishing a testbed, we could provide a platform for other researchers within our group to conduct their research and experiments.

>>>>

As previously discussed, this dissertation makes two key contributions to the literature surrounding performance evaluation of Edge Computing environments.
These contributions, together with ancillary contributions and results, will be summarized in the two sections presented in this chapter.

In \cref{summary:methodology}, we discuss our methodology for evaluating the performance of latency-sensitive, feedback-loop-based applications enabled by Edge Computing.
This methodology involves emulating only the end-user component of such applications while using the real network and compute components of the system. 
We further discuss the validation of this approach through two representative use-cases for \gls{XR} and \glspl{CPS}.

Next, in \cref{summary:realism}, we explore the extent to which our client-side emulation of a latency-sensitive, feedback-loop application on Edge Computing needs to capture certain details of the physical world.
We broach this question through the elaboration of a detailed model of human behavior in \gls{WCA}, which we then employ in our methodology, in conjunction with a mathematical framework, to study the the potential for optimization of \gls{WCA} applications.

@@

This thesis makes a contribution to the field of performance evaluation of Edge Computing systems by investigating the applications of a methodology for the study of latency-sensitive applications deployed on Edge environments.
These applications, such as \gls{WCA} and \glspl{NCS}, are challenging to study and scale due to their cyber-physical nature, complexity, and requirements for low latency and high reliability.

The main focus of this work is to enhance the accuracy and realism of results related to Edge infrastructure.
In particular, the proposed methodology aims to provide a more comprehensive and realistic assessment of the performance of these latency-sensitive applications with respect to previous approaches, allowing for a better understanding of the strengths and limitations of Edge Computing.
Our results will contribute to the development of new techniques and approaches for improving the performance and reliability of Edge Computing systems.

@@

Our emulated methodology presents several advantages over alternatives, simulated or fully-practical approaches.
To begin with, and perhaps most importantly, it makes our approach portable and applicable to real Edge computing deployments.
By limiting the emulation to the workload, our approach can be directly employed on pre-existing Edge computing infrastructure in order to benchmark applications before they are deployed.
This makes our approach valuable to not only to the research community, but also to system designers and application developers.

@@

\medskip
In the following subsections we will discuss two illustrative use-cases for our methodology; \acl{WCA} and \acl{NCS} research and benchmarking.
\Cref{summary:methodology:usecase_wca} discusses the application of our methodology to \gls{WCA}.
This work was originally published in \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid}.
Next, in \cref{summary:methodology:usecase_ncs} we expand this discussion to \glspl{NCS}, a discussion originally approached in \cref{paper:olguinmunoz2022cleave}.

\subsection{Use-case: \acl{WCA}}\label{summary:methodology:usecase_wca}

@@

\subsection{Use-case: \acsp{NCS}}\label{summary:methodology:usecase_ncs}

We discuss an initial use-case study of our methodology in the context of \glspl{NCS} and similar \glspl{CPS} in \cref{paper:olguinmunoz2022cleave}.
In this work, we present \gls{CLEAVE}, a software-based framework for scalable and repeatable benchmarking of Edge-native \aclp{NCS}.
The growing adoption of Edge computing has made it important to understand the strengths and weaknesses of different Edge concepts for the deployment of such \glspl{CPS}.

@@

After our initial implementation and experimentation with our methodology for \gls{WCA} applications, we arrived at the realization that the creation of a dedicated testbed for experimental Edge Computing research would greatly augment the scope and impact of our work.
Establishing such testbed would provide a dedicated and controlled environment that can accelerate future research by eliminating the need for ad-hoc implementations for each new study.
Moreover, having a standardized baseline would make it easier to compare results across different studies.
The testbed could serve as a platform for researchers to test new ideas and compare them to the results of previous experiments.
This would enable more efficient and accurate assessments of the methodology's performance and facilitate the development of new methods and techniques.
Additionally, by establishing a testbed, we could provide a platform for other researchers within our group to conduct their research and experiments.

### 2023-04-22 body/part1/chapter3-contributions/wca-implications.tex
contrib tree

\section{Implications of the methodology in the context of \acs{WCA}}\label{sec:extendingmethodology}

In this section, we aim to delve deeper into the implications of our methodology for the optimization of \gls{WCA}.
To this end, we must first address a limitation of our current model of human behavior in the context of these applications.
Our work in \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid} has used a model that emulates a rigid and unyielding human who makes no mistakes and is not affected by low responsiveness in the system.
However, this is not an accurate representation of real human behavior, and a more realistic human model must take into consideration the perceived feedback latency and its effects on human reactions.

To build a more accurate model, we must first characterize human behavior in the context of \gls{WCA}.
We achieve this through an experimental study measuring human performance in various scenarios, taking into account the impact of latency in the system to human behavior.
The results of this study are then used to build a more representative model of human behavior, which we further use to provide a foundation for the optimization of these applications.

Once we have established more accurate model of human behavior, we can then begin to explore the implications of our methodology for the optimization of \gls{WCA}.
This includes evaluating the impact of various optimization strategies on human performance and identifying opportunities for improvement.
By incorporating a realistic model of human behavior, we will be able to build a more comprehensive understanding of the challenges involved in optimizing \gls{WCA} and how our methodology can be applied to address these challenges.

\subsection{Characterization of \glsfmtshort{WCA}}
\todo[inline]{Rework}

@@

\subsubsection{Implications}

@@

\cref{paper:olguinmunoz2023realistic} discusses the elaboration of a model of human behavior for \gls{WCA}.
Built upon the data and results from \cref{paper:olguinmunoz2021impact}, our model is capable of dynamically generating realistic predictions of execution times for steps in a \gls{WCA} based on historical system state data and a parameterized level of neuroticism.
Furthermore, we implement a new version of the EdgeDroid tool capable of using this new model, together with a novel procedure for the generation of realistic synthetic traces of frames for \gls{WCA}, to fully emulate a human in the context of these systems.

>>>>

\section{Exploring the need for realism in client-side emulation}\label{summary:realism}

In this section, we aim to delve deeper into the implications of our methodology for the optimization of \gls{WCA}.
In particular, we wish to explore the impact improving the realism of the client-side emulation has on the potential for optimization of the applications under evaluation.
We approach this question by developing an improved, more detailed version of our model of human behavior in \gls{WCA} used in the EdgeDroid tool, which we then incorporate into our methodology along with a mathematical framework to study the potential for optimizing \gls{WCA} applications.

Our work in \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid} has used a model that emulates a rigid and unyielding human who makes no mistakes and is not affected by low responsiveness in the system.
However, this is not an accurate representation of real human behavior, and a more realistic human model must take into consideration the perceived feedback latency and its effects on human reactions.
In order to develop a more accurate human model in the context of \gls{WCA}, we first conduct an experimental study that measures human performance in different scenarios, taking into account the impact of system latency on the behavior of humans.
The results of this study serve as the foundation for building a more representative model of human behavior.

Once we have established a more accurate model of human behavior, we can explore its implications for the optimization of \gls{WCA} through our methodology.
By improving the realism of our model of human behavior, we aim to gain novel understanding of the challenges involved in optimizing \gls{WCA} and how our methodology can be effectively applied to address these.
This research highlights the importance of the realism of emulations as  crucial factor in the application of our methodology.

\subsection{Definitions relating to the operation of \glsfmtshort{WCA} applications}

@@

\subsubsection{Implications for \glsfmtshort{WCA}}

@@

\cref{paper:olguinmunoz2023realistic} discusses the elaboration of a model of human behavior for \gls{WCA}.
Built upon the data and results from \cref{paper:olguinmunoz2021impact}, our model is capable of dynamically generating realistic predictions of execution times for steps in a \gls{WCA} based on historical system state data and a parameterized level of neuroticism.
Furthermore, we implement a new version of the EdgeDroid tool capable of using this new model, together with a novel procedure for the generation of realistic synthetic traces of frames for \gls{WCA}, to provide a fully end-to-end emulation of a human in the context of these systems.
